---
title: "Main text statistics for Luminex Recommendation Paper"
author: "Forrest Jones"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,fig.height = 7,
                      fig.width = 10)
```


```{r functions, eval=TRUE}

source("source/final_code/luminex_recommendation/R/packages.R")
source("source/final_code/shared/utils.R")


```


```{r data, eval=TRUE}

source("source/final_code/luminex_recommendation/R/load-data.R")

case_days <- getTimingData()
wideData <- demo_id <-getWideData() %>% mutate(blood=factor(blood))

```

All values calculated here are not found in the main figures or supplement.

## Methods

### Study Population 

```{r timingstuff}


# #use this code to eventually replace sample timing things
# 
# #limit luminex data to just id, day, sample
# Luminex <-getLuminexSamplesRAU() %>%
#             left_join(wideData) %>%
#             filter(status==c("Case","Contact")) %>%
#             distinct(id,day,sample)
# 
# pic_DATES <- bind_rows(
# 
# # file 1
# read.csv(here::here('data','raw_data','cohort_data','dates_data','pic_date_OCT.csv'),as.is=T)%>%
#         select(SID,starts_with("CDAT"))  %>%
#         gather(day,date_visit,-SID) %>%
#         mutate(day=str_remove(day,"CDAT") )%>%
#         mutate(day=case_when(
#                 day =="7D" ~ 7,
#                 day =="1M" ~ 30,
#                 day =="3M" ~ 90,
#                 day =="6M" ~ 180,
#                 day =="9M" ~ 270,
#                 day =="12M" ~ 360)) %>%
#                 filter(date_visit!="")%>%
#         mutate(date_visit=as.Date(date_visit,
#                                    format="%d-%b-%Y",origin="1970-1-1")),
# 
# 
# # #file 2
# read.csv(here::here('data','raw_data','cohort_data','dates_data','pic_final_2012_FromAshraf_6.2015.csv'),as.is=T)%>%
#         select(sid,starts_with("CDAT")) %>%
#         rename(SID=sid) %>%
#         gather(day,date_visit,-SID) %>%
#         mutate(day=str_remove(day,"cdat") %>% toupper) %>%
#         filter(day %in% c("7D","1M","3M","6M",
#                           "9M","12M"
#                           ))%>%
#         mutate(day=case_when(
#                 day =="7D" ~ 7,
#                 day =="1M" ~ 30,
#                 day =="3M" ~ 90,
#                 day =="6M" ~ 180,
#                 day =="9M" ~ 270,
#                 day =="12M" ~ 360)) %>%
#                 filter(date_visit!="")%>%
#         mutate(date_visit=as.Date(date_visit,
#                            format="%d/%m/%Y",origin="1970-1-1")) 
# 
# )  %>%
#         mutate(id=paste0("P",SID)) %>%
#         mutate(sample=paste0(id," d",day)) %>%
#         distinct(id,day,sample,date_visit)
# 
# 
# ## to make it easier to get into same format at smic_lab, I will deal with the three visits
# ## seperatley. Not efficient but...
# rename_smic_lab <- function(dat,day=7){
#     dat  %>% rename_at(.vars = vars(ends_with(day)),
#                              .funs = funs(sub(paste0(day,"$"), "", .))) %>%
#     rename_all(toupper)
# 
# }
# 
# ## merge with long term dates for cases
# path_to_excel_file <- here('data','raw_data','cohort_data','dates_data','SMIC follow up Enrollment Dates.xlsx')
# 
# smic_post365_dates <- full_join(
#     read_excel(path_to_excel_file,sheet = "D540",na=c(NA,"Dropout","--")) %>% rename(SID = `SMIC ID`,
#                                                                                                               date_540d =  `Date of Enrollment, D540`),
#     read_excel(path_to_excel_file,sheet = "D720",na=c(NA,"Dropout","--")) %>% rename(SID = `SMIC ID`,
#                                                                                                               date_720d =  `Date of Enrollment, D720`)) %>% 
# full_join(.,
#           read_excel(path_to_excel_file,sheet = "D900",na=c(NA,"Dropout","--")) %>% rename(SID = `SMIC ID`,
#                                                                                                                     date_900d =  `Date of Enrollment, D900`)
#           ) %>%
# mutate(date_540d = as.Date(date_540d,format = "%d/%m/%y"),
#        date_720d = as.Date(date_720d,format = "%d/%m/%y"),
#        date_900d = as.Date(date_900d,format = "%d/%m/%y"),
#        date_540d = if_else(SID == 29,as.Date("2013-10-01"),date_540d) ## fixing this seemingly clear mistake
#        ) %>%
# filter(!(is.na(date_540d) & is.na(date_720d) & is.na(date_900d)))
# 
# 
# smic_lab2_540 <- read.csv(here('data','raw_data','cohort_data','dates_data','SMICdata.vbx.plasma.Azman.2018.csv')) %>%
# select(sid,matches("7$")) %>%
# rename_smic_lab(day="7") %>%
# mutate(LBDAY = 540) %>%
# select(-starts_with("pm"),-starts_with("sm")) ## getting rid of IgM since we bring in seperateley
# 
# 
# 
# smic_DATES<- bind_rows(
#         
#         #file 3
#         read.csv(here('data','raw_data','cohort_data','dates_data','smic_date_OCT.csv'),as.is=T)%>%
#                 select(SID,starts_with("CDAT"))  %>%
#                 gather(day,date_visit,-SID) %>%
#                 mutate(day=str_remove(day,"CDAT") )%>%
#                 mutate(day=case_when(
#                         day =="7D" ~ 7,
#                         day =="1M" ~ 30,
#                         day =="3M" ~ 90,
#                         day =="6M" ~ 180,
#                         day =="9M" ~ 270,
#                         day =="12M" ~ 360))%>%
#                         filter(date_visit!="")%>%
#         mutate(date_visit=as.Date(date_visit,
#                                    format="%d-%b-%Y",origin="1970-1-1")),
#         
#         
#         
#         #file 4
#         read.csv(here('data','raw_data','cohort_data','dates_data','SMICdata.vbx.plasma.Azman.2018.csv')) %>%
#                 select(SID=sid,starts_with("CDAT"))%>%
#                 mutate(date_visit=as.character(cdat7d)) %>%
#                 select(-cdat7d) %>%
#                 mutate(day=7) %>%
#                 filter(date_visit!="") %>%
#         mutate(date_visit=as.Date(date_visit,
#                                    format="%d-%b-%y",origin="1970-1-1")),
# 
# 
#         #file 5
#         smic_post365_dates %>%
#                 gather(day,date_visit,-SID) %>%
#                 mutate(day=str_remove(day,"date_")) %>%
#                 mutate(day=str_remove(day,"d") %>% as.numeric) %>%
#                 filter(!is.na(date_visit)),
# 
# 
# #         #file 6
#         smic_lab2_540 %>% select(SID,date_visit=QFHDAT,day=LBDAY) %>%
#                 filter(date_visit!="") %>%
#         mutate(date_visit=as.Date(date_visit,
#                                    format="%d-%b-%y",origin="1970-1-1"))
# ) %>% mutate(id=paste0("S",SID)) %>%
#         mutate(sample=paste0(id," d",day)) %>%
#         distinct(id,day,sample,date_visit)
# 
# 
# theDATES <- bind_rows(smic_DATES,pic_DATES)
# 
# master_data <- readstata13::read.dta13("data/raw_data/cohort_data/CIRS PIC SMIC Master DB March 2019.dta") 
# labels <- readstata13::varlabel(master_data)
# labels <- data.frame(variable=names(labels),description=as.character(labels))
# 
# master_data_ltd<- filter(master_data, studycode %in% c("SMIC","PIC")) %>%
#     mutate(studycode=substr(studycode,1,1)) %>%
#     mutate(id=paste0(studycode,sid)) %>%
#     select(id, date_enroll=dten,diadur) %>%
#     filter(id %in% Luminex$id)
# 
# 
# huh <- distinct(Luminex,id,sample,day)  %>%
#                 left_join(master_data_ltd) %>% 
#                 left_join(theDATES) %>%
#                 mutate(diff=(date_visit-date_enroll)%>% as.numeric) %>%
#                 mutate(diff=ifelse(is.na(diff),day,diff)) %>%
#                 # - diarrhea duration
#                 # - incubation period (1.4 days)
#                 dplyr::mutate(symptom_lag = ifelse(is.na(diadur),0,diadur/24)) %>%
#                 #calculate time to enrollment
#                 mutate(tte=round(symptom_lag+1.4)) %>%
#                 mutate(new_actual=tte+diff)
# 
# 
# filter(huh, day %in% c(2,7)) %>%
#         ggplot(aes(x=new_actual))+
#                 geom_histogram(aes(fill=factor(day)))


```

```{r study-pop}
 
demo_id %>% filter(studycode  %in% c("SMIC","PIC")) %>% pull(date_enroll) %>%
        lubridate::year() %>%
        summary()


#those tested after checking inventory
filter(analysisData$compare_data, !str_detect(batch,"Vaccine"))  %>%
                distinct(id,sample,status,age) %>%
                group_by(age<10) %>%
                summarize(samples=n(),
                          length(unique(id))) 
                

filter(analysisData$compare_data, str_detect(batch,"Vaccine"))  %>%
                distinct(id,sample,status,age) %>%
                summarize(samples=n(),
                          min(age),
                          max(age),
                          length(unique(id))) 




```

### Serological testing and data processing 


```{r testing}

analysisData$compare_data$date %>% summary()

```

### Statistical Analysis

```{r analysis}

raw <- read_rds(paste0("data/generated_data/2021-10-25-raw-data.rds"))

raw %>% 
  filter(subclass=="total") %>%
  mutate(lowCount=Count<30) %>%
  pull(lowCount)%>%
  mean()%>%
  round(4)*100 



```

## Results


### Description of individuals and timing of samples


```{r describe-sample}

characteristics_sample <- casecon_analysis %>% 
            distinct(id,sample,day,day_actual,age,age_group,sex,blood,culture,
                     status,dehydration, `Case Watery Diarrhea Day 7`,
                     `Contact Watery Diarrhea Day 2`,hospital_hrs)


characteristics_sample %>% group_by(status) %>% summarize(n=n(),
                                                          min(day_actual),
                                                          max(day_actual)
                                                          ) 

characteristics_id <- casecon_analysis %>% 
            distinct(id,age,age_group,sex,blood,culture,
                     status,dehydration, `Case Watery Diarrhea Day 7`,
                     `Contact Watery Diarrhea Day 2`,hospital_hrs)



table(characteristics_id$culture)/48 

characteristics_id %>% group_by(status) %>%
          summarize(
            round(median(age)), round(quantile(age,0.25)),
            round(quantile(age,0.75)),
            
            male=round(mean(sex=="male"),2),
            Oblood=round(mean(blood=="O"),2)
            
          )

# 
# characteristics_id %>% filter(status=="Case") %>%
#           summarize(
#             round(median(hospital_hrs)), round(quantile(hospital_hrs,0.25)),
#             round(quantile(hospital_hrs,0.75))
#           )



characteristics_sample %>% filter(status=="Case") %>% 
      group_by(day) %>%
      summarise(
                n=n(),
                percent=n/48,
                min(day_actual),
                max(day_actual)
                )


filter(characteristics_sample,day>=270) %>%
        group_by(id) %>%
        summarize(samples=n()) %>%
        group_by(samples) %>%
        summarize(n=n())

age_group_ltfu <- characteristics_sample %>% filter(status=="Case") %>% 
      group_by(day,age_group) %>%
      summarise(
                n=n()
                ) 


age_group_ltfu %>% group_by(age_group) %>%
      mutate(n=n/max(n)) %>%
      spread(age_group,n) 


# age_group_ltfu %>% group_by(age_group) %>%
#       # mutate(n=n/max(n)) %>%
#       spread(age_group,n) %>%
#       filter(!is.na(`<5 years`)) %>% select(-day) %>%
#       as.matrix() %>%
#       chisq.test()
 


age_group_ltfu %>% ggplot(aes(x=as.factor(as.numeric(day)),y=n,col=age_group))+
                  geom_point()+
                  geom_line(aes(group=age_group))




```


### Kinetics of biomarkers in confirmed cholera cases 



```{r kinetics}

# Among CT-B,CT-H, Inaba OSP and Ogawa OSP, the median day of peak levels ranged between 25-36 days for IgG, at 8 days for IgA, and between 8 - 63.7 days for IgM. Across isotypes for Ogawa and Inaba OSP,  individual variability in MBA measurements 

casecon_analysis %>% 
          filter(antigen %in% c("OgawaOSPBSA","InabaOSPBSA","CtxB","CTHT")) %>%
          filter(status=="Case") %>%
          filter(str_detect(full_test,"RAU")) %>%
          group_by(id,isotype,full_test) %>%
              filter(value==max(value)) %>%
              summarize(day=mean(day_actual)) %>%
          group_by(full_test,isotype) %>%
          summarize(n(),
                    day=round(median(day))
                    ) %>%
          arrange(isotype,day) %>%
          group_by(isotype) %>%
          summarize(min=min(day),
                    max=max(day))


# Across isotypes for Ogawa and Inaba OSP,  individual variability in MBA measurements increased at day 7 and did not return to baseline levels until over 180 days (Figure 2); in contrast, individual variability remained relatively constant over time for CT-B and CT-H.
          
casecon_analysis %>% 
          filter(antigen %in% c("OgawaOSPBSA","InabaOSPBSA","CtxB","CTHT")) %>%
          filter(status=="Case") %>%
          filter(str_detect(full_test,"RAU")) %>%
          filter(!day %in% c(720,1080)) %>%
              group_by(day,isotype,antigen) %>%
                summarize(sd=sd(value)) %>% 
            ggplot(aes(x=factor(day),y=sd,col=isotype))+
              geom_point()+
              geom_line(aes(group=isotype))+
              facet_wrap(.~antigen)



```


```{r}

# summary_cov_df <- read_rds(
#            paste0("source/final_code/luminex_recommendation/generated_rds/decay_model_results/covariates/summary_cov_df.rds"))
# 
# 
# #Fold change difference table
# FC_diff_table<- summary_cov_df %>% filter(str_detect(parameter,"baseline_ratio|foldchange_ratio")) %>%
#                 select(marker,covariate,parameter,`X2.5.`,`X50.`,`X97.5.`) %>%
#                 gather(quantile,value,-c(marker,covariate,parameter)) %>%
#                 # mutate(value=10^value) %>%
#                 spread(quantile,value) %>% 
#                 mutate(CI =paste0(signif(`X50.`,3)," (",
#                        signif(`X2.5.`,3),"-",
#                        signif(`X97.5.`,3),")")) %>%
#                 mutate(CI=ifelse((`X2.5.`<1)+(`X97.5.`>1)==2,CI,paste0(CI,"*"))) 
# 
# 
# 
# 
# #half life change difference table
# HL_diff_table <- summary_cov_df %>% filter(str_detect(parameter,"halflife_diff")) %>%
#                 select(marker,covariate,parameter,`X2.5.`,`X50.`,`X97.5.`) %>%
#                 mutate(CI =paste0(round(`X50.`)," (",
#                        round(`X2.5.`),"-",
#                        round(`X97.5.`),")")) %>%
#                 mutate(CI=ifelse(`X2.5.`*`X97.5.`>0,paste0(CI,"*"),CI)) 
# 
# 
# # Under 10 differences
# FC_diff_table %>% filter(covariate=="Under 10") %>%
#         filter(marker %in% c("Luminex_IgG_OgawaOSPBSA","Luminex_IgG_InabaOSPBSA")) %>%
#         filter(parameter=="foldchange_ratio") %>%
#         mutate(`X2.5.`=round(1/`X2.5.`,1),
#                `X50.`=round(1/`X50.`,1),
#                `X97.5.`=round(1/`X97.5.`,1)
#                )
# 
# 
# HL_diff_table %>% filter(covariate=="Under 10") %>%
#         filter(marker %in% c("Luminex_IgG_OgawaOSPBSA","Luminex_IgG_InabaOSPBSA")) 
# 
# 
# # Ogawa vs. Inabadifferences
# FC_diff_table %>% filter(covariate=="Ogawa") %>%
#         filter(marker %in% c("Luminex_IgM_OgawaOSPBSA",
#                              "Luminex_IgG_OgawaOSPBSA",
#                              "Luminex_IgA_OgawaOSPBSA"
#                              )) %>%
#         filter(parameter=="foldchange_ratio")  %>%
#         mutate(`X2.5.`=round(`X2.5.`,1),
#                `X50.`=round(`X50.`,1),
#                `X97.5.`=round(`X97.5.`,1)
#                )
# 
# FC_diff_table %>% filter(covariate=="Ogawa") %>%
#         filter(marker %in% c("Luminex_IgM_InabaOSPBSA",
#                              "Luminex_IgG_InabaOSPBSA",
#                              "Luminex_IgA_InabaOSPBSA"
#                              )) %>%
#         filter(parameter=="foldchange_ratio") %>%
#         mutate(`X2.5.`=round(1/`X2.5.`,1),
#                `X50.`=round(1/`X50.`,1),
#                `X97.5.`=round(1/`X97.5.`,1)
#                )


```



### Identification of recent infections with cross-sectional serologic measurements 

```{r tvCVAUC}

cvAUCdf <- read_rds("source/final_code/luminex_recommendation/generated_rds/random_forest_results/cvAUC_importance/1_scan_cvAUC_df.rds")

cvAUCdf %>% 
        filter(predictors=="Net MFI All") %>% 
        pull(cvAUC) %>% summary()


```

```{r cvAUC, eval=FALSE}

#get ratios to to see relative cvAUC of traditional and MBA markers
marker_set_names <- c( "Vibriocidal &\nELISA Markers","Vibriocidal\nMarkers","ELISA\nMarkers","All MBA\nMarkers","IgG MBA\nMarkers", "IgA MBA\nMarkers","IgM MBA\nMarkers")

marker_cvAUC_list <- read_rds(
           "source/final_code/luminex_recommendation/generated_rds/random_forest_results/cvAUC_importance/cvAUC_list.rds"
          )  



cv_marker_df <- data.frame()
times <- c(45,120,200,300)

for(k in 1:length(marker_cvAUC_list)){
  for(j in as.character(times)){
        cv_marker_df <- bind_rows(cv_marker_df,
                                   marker_cvAUC_list[[k]][[j]][[4]] %>%
                                          mutate(
                                  model=marker_set_names[k])
                                  )
  }
}

cv_marker_df %>% filter(model %in% c( "Vibriocidal &\nELISA Markers","All MBA\nMarkers")) %>%
                dplyr::select(model,end_window,cvAUC) %>%
                mutate(model=substr(model,1,7)) %>%
                spread(model,cvAUC) %>%
                mutate(ratio=`All MBA`/Vibrioc)


#get ratios to to see relative cvAUC of traditional and MBA markers
cv_marker_df %>% filter(model %in% c( "Vibriocidal &\nELISA Markers","All MBA\nMarkers")) %>%
                dplyr::select(model,end_window,cvAUC) %>%
                mutate(model=substr(model,1,7)) %>%
                spread(model,cvAUC) %>%
                mutate(ratio=`All MBA`/Vibrioc)

#get relative reduction
cv_marker_df %>% filter(model %in% marker_set_names[str_detect(marker_set_names,"Ig[A|G|M]")]) %>%
                dplyr::select(model,end_window,cvAUC) %>%
                left_join(cv_marker_df %>%
                                  filter(model %in% c("All MBA\nMarkers")) %>%
                                  dplyr::select(end_window,AllMBA=cvAUC) ,
                          by = "end_window"
                                  ) %>%
                mutate(reduction=round((AllMBA-cvAUC)*100)) %>%
                mutate(relative_reduction=round((AllMBA-cvAUC)/AllMBA*100)) %>%
                filter(end_window==200)

        

```


```{r sens-spec}

tvsens_df <-read_rds("source/final_code/luminex_recommendation/generated_rds/random_forest_results/tvsens-spec/tvsens_df.rds")

tvsens_df %>% filter(seropos_type=="spec90_seropos") %>%
                group_by(end_window) %>%
                summarize(min(median) %>% round(2),
                          max(median) %>% round(2)
                          )

fit_ssy_df<- read_rds("source/final_code/luminex_recommendation/generated_rds/random_forest_results/tvsens-spec/cutoff_fit_ssy_df.rds")
# can use this to find the sensitivity and specificity for your cutoff here
fit_ssy_df %>% filter(round(Luminex_IgG_CtxB,1)==-3.1,
                      round(Luminex_IgG_OgawaOSPBSA,1)==-4
                       )
```

## Discussion

